# Results

```{r}
library(tidyverse)
library(stringr)
library(plotly)
library(shiny)
library(shinydashboard)
library(ggrepel)
library(knitr)
library(gridExtra)
```


## Popular movies vs Oscar Winners 

```{r}
#manipulating tables
bo_oscar_highest_data <- left_join(oscar_bo, highest_grossing, by="Year") %>% select(Year, Overall_Movie, Overall_Boxoffice, Oscar_Movie, Oscar_Boxoffice)

rows <- nrow(bo_oscar_highest_data)

bo_oscar_temp <- bo_oscar_highest_data %>% select(Year, Oscar_Boxoffice, Oscar_Movie) %>% rename(Boxoffice = Oscar_Boxoffice, Movie = Oscar_Movie) %>% mutate(Group = rep("Oscar", rows))
bo_overall_temp <- bo_oscar_highest_data %>% select(Year, Overall_Boxoffice, Overall_Movie) %>% rename(Boxoffice = Overall_Boxoffice, Movie = Overall_Movie) %>% mutate(Group = rep("Overall", rows))

bo_oscar_highest_data <- rbind(bo_oscar_temp, bo_overall_temp)
```

```{r}
#plotting
# bo_oscar_highest_data %>% 
#   ggplot(aes(x = Year, y = Boxoffice, col = Group)) + 
#   geom_line(size = 1) +
#   scale_y_continuous(labels = scales::comma) +
#   theme_light() +
#   scale_color_manual(values = c("#FAC983", "#000000")) +
#   theme(legend.title= element_blank())


plot_ly() %>%
  add_lines(x = bo_oscar_temp$Year, y = bo_oscar_temp$Boxoffice, 
            color = I("#FAC983"), 
            name = "Oscar Best Picture",
            text = bo_oscar_temp$Movie,
            hoverinfo = 'text') %>%
  add_lines(x = bo_oscar_temp$Year, y = bo_overall_temp$Boxoffice, 
            color = I("black"), 
            name = "Highest Grossest Film",
            text = bo_overall_temp$Movie,
            hoverinfo = 'text') %>%
  layout(legend = list(x = 100, y = 0.5), title = 'Best Picture vs Highest Grossing Boxoffice')
```


## IMDB Ratings of Oscar Movies over Time

```{r}
# Dashboard ----------------------------------------------------------------
ui <- dashboardPage(
  dashboardHeader(title = "Test"),
  dashboardSidebar(),
  dashboardBody(tags$head(tags$style(
    HTML("img.small-img {
          max-width: 75px;
          }")
  )),
  plotlyOutput("hoverplot"))
)

server <- function(input, output, session) {
  output$hoverplot <- renderPlotly({
    plot_ly(
      imdb_oscar,
      x         = ~ Year,
      y         = ~ Rating,
      type      = 'scatter',
      mode      = 'markers',
      hoverinfo = 'none',
      source = "hoverplotsource",
      customdata = ~ Image
    ) %>%
      event_register('plotly_hover') %>%
      event_register('plotly_unhover')
  })

  hover_event <- reactive({
    event_data(event = "plotly_hover", source = "hoverplotsource")
  })

  unhover_event <- reactive({
    event_data(event = "plotly_unhover", source = "hoverplotsource")
  })

  hoverplotlyProxy <- plotlyProxy("hoverplot", session)

  observeEvent(unhover_event(), {
    hoverplotlyProxy %>%
      plotlyProxyInvoke("relayout", list(images = list(NULL)))
  })

  observeEvent(hover_event(), {
    hoverplotlyProxy %>%
      plotlyProxyInvoke("relayout", list(images = list(
        list(
          source = hover_event()$customdata,
          xref = "x",
          yref = "y",
          x = hover_event()$x,
          y = hover_event()$y,
          sizex = 20,
          sizey = 20,
          opacity = 1
        )
      )))
  })
}

shinyApp(ui = ui, server = server)

#include_app(t, height = "400px")
```

```{r}
theme_oscars = function(base_size = 12, grid_width = 0.2) {
  bg_rect = element_rect(fill = "#EBAB47", color = "#EBAB47")

  theme_bw(base_size) +
    theme(
          plot.background = bg_rect,
          legend.background = bg_rect,
          panel.background = bg_rect,
          panel.grid.major = element_line(colour = "#3c3030", size = grid_width),
          panel.grid.minor = element_line(colour = "#3c3030", size = grid_width),
          legend.key.width = unit(1.5, "line"),
          legend.key = element_blank(),
          axis.title.x = element_text(margin = unit(c(0.5, 0.25, 0.5, 0.25), "line")),
          axis.title.y = element_text(margin = unit(c(0, 1, 0, 0.1), "line")),
          axis.ticks = element_blank(),
          panel.border = element_blank())
}
```


## Total Nominations and Wins of actors

```{r}
#data cleaning
all_actors <- data.frame(Actors = c(actor_lead$Actor, actor_sup$Actor))
all_actors_filter <- as.data.frame(table(all_actors))
all_actors_filter <- all_actors_filter %>% filter(Freq > 4) %>% arrange(all_actors)
actor_wins <- actor_wins %>% arrange(Name)
all_actors_filter <- all_actors_filter %>% mutate(Noms = all_actors_filter$Freq - actor_wins$Wins)

barplot_data <- function(actor, times, status) {
  
  new_df <- data.frame(Actor=character(0), Status=character(0))
  
  n <- length(actor)
  for(i in 1:n) {
  actor_rep <- rep(actor[i], times[i])
  status_rep <- rep(status, times[i])
  new_df <- rbind(new_df, data.frame(Actor = actor_rep, Status = status_rep))
  }
  return(new_df)
}

actor_wins_plot <- barplot_data(actor_wins$Name, actor_wins$Wins, "W")
actor_noms_plot <- barplot_data(all_actors_filter$all_actors, all_actors_filter$Noms, "N")
all_actors_new <- rbind(actor_wins_plot, actor_noms_plot)
all_actors_new <- all_actors_new %>% mutate(Status = as.factor(Status))

all_actors <- all_actors %>% filter(Actors %in% all_actors_filter$all_actors)
```

```{r}
#new plotting
all_actors_new %>% 
  ggplot(aes(x=fct_rev(fct_infreq(Actor)))) + 
  geom_bar(aes(fill=forcats::fct_rev(Status))) + 
  coord_flip() +
  ylab("Number of Oscar Nominations") +
  theme_oscars(base_size = 10, grid_width = 0.4) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(hjust = 1.38, margin = unit(c(0.25, 0.25, 2, 0.25), "line")),
    axis.text.y = element_text(size = 10, margin = unit(c(0, -1, 0, 0), "line")),
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 10),
    legend.position = "none"
  ) + scale_fill_manual(values = c("#33a02c", "#000000")) +
  scale_y_continuous(breaks=c(0,2, 4, 6,8, 10, 12))
```




## Total Nominations and Wins of actresses

```{r}
#data cleaning
all_actress <- data.frame(Actress = c(actress_lead$Actress, actress_sup$Actress))
all_actress_filter <- as.data.frame(table(all_actress))

all_actress_filter <- all_actress_filter %>% filter(Freq > 4) %>% arrange(all_actress)
actress_wins <- actress_wins %>% arrange(Name)
all_actress_filter <- all_actress_filter %>% mutate(Noms = all_actress_filter$Freq - actress_wins$Wins)

actress_wins_plot <- barplot_data(actress_wins$Name, actress_wins$Wins, "W")
actress_noms_plot <- barplot_data(all_actress_filter$all_actress, all_actress_filter$Noms, "N")
all_actress_new <- rbind(actress_wins_plot, actress_noms_plot)
all_actress_new <- all_actress_new %>% mutate(Status = as.factor(Status))
```

```{r}
#new plotting
all_actress_new %>% 
  ggplot(aes(x=fct_rev(fct_infreq(Actor)))) + 
  geom_bar(aes(fill=forcats::fct_rev(Status))) + 
  coord_flip() +
  ylab("Number of Oscar Nominations") +
  theme_oscars(base_size = 10, grid_width = 0.4) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(hjust = 1.38, margin = unit(c(0.25, 0.25, 2, 0.25), "line")),
    axis.text.y = element_text(size = 10, margin = unit(c(0, -1, 0, 0), "line")),
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 10),
    legend.position = "none"
  ) + scale_fill_manual(values = c("#33a02c", "#000000"))
```


```{r}
#plotting
all_actress %>% 
  ggplot(aes(x=fct_rev(fct_infreq(Actress)))) + 
  geom_bar(fill="black") + 
  coord_flip() +
  ylab("Number of Oscar Nominations") +
  theme_oscars(base_size = 10, grid_width = 0.4) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(size = 15),
    plot.subtitle = element_text(hjust = 1.38, margin = unit(c(0.25, 0.25, 2, 0.25), "line")),
    axis.text.y = element_text(size = 10, margin = unit(c(0, -1, 0, 0), "line")),
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 10),
    legend.position = "none"
  )
  #ggtitle("Number of time Actors were Nominated for an Oscar")
```


## Female-lead Movie Box Office vs Other Nominees that Year

```{r}
female_nom_bo %>% mutate(Status = ifelse(Status == "N", "Male-lead Nominee", "Female-lead Winner")) %>%
  ggplot(aes(x=as.factor(Year), y=Boxoffice, col=as.factor(Status)), label=Movie) +
  geom_point(size=2) +
  ylab("Boxoffice Earnings (in Millions)") +
  theme_light() +
  #geom_text(aes(label = ifelse(Movie %in% female_winners$Movie, Movie, "")),hjust=0, vjust=0, position = position_dodge()) +
  scale_color_manual(values = c("#000000", "#EBAB47")) +
  theme(axis.title.x = element_blank(), legend.title= element_blank(),
        legend.position = c(0.15, 0.9)) +
  geom_label_repel(aes(label = ifelse(Movie %in% female_winners$Movie, Movie, "")), size = 3, max.overlaps = Inf) +
  ggtitle("Comparing the Boxoffice earnings of female-lead Oscar Best Picture winners\n with the other nominees of that year")
```


## Average Rating of Oscar Nominees over Time

```{r}
imdb_oscar %>% group_by(Year) %>% 
  summarise(average_rating = mean(Rating)) %>%
  ggplot(aes(x=Year, y=average_rating)) + geom_line() + geom_point(col = "#ea2c0b") +
  ggtitle("Average audience rating of Oscar Best Picture nominees over the years") +
  ylab("Average Rating (out of 10)") +
  theme_oscars(base_size = 10, grid_width = 0.4)
```

## Average Runtime of Oscar Nominees over Time

```{r}
runtimePattern <- "\\smin"
imdb_oscar %>% 
  mutate(Time = as.numeric(str_replace(Time, runtimePattern, ""))) %>% 
  group_by(Year) %>% 
  summarise(average_runtime = mean(as.numeric(Time))) %>%
  ggplot(aes(x=Year, y=average_runtime)) + geom_line() + geom_point(col = "#ea2c0b") +
  ggtitle("Average runtime of Oscar Best Picture nominees over the years") +
  ylab("Average Runtime (mins)") +
  theme_oscars(base_size = 10, grid_width = 0.4)
```


## Average Runtime vs Rating of Oscar Best Picture Nominees

```{r}
runtimePattern <- "\\smin"
imdb_oscar %>% 
  mutate(Time = as.numeric(str_replace(Time, runtimePattern, ""))) %>%
  ggplot(aes(x=Time, y=Rating)) + geom_point(col = "#ea2c0b") +
  ggtitle("Audience Rating vs Runtime of Oscar Nominated Pictures") +
  xlab("Runtime (mins)") + ylab("Rating (out of 10)") +
  theme_oscars(base_size = 10, grid_width = 0.4)
```


## Best Lead Actors Race over Time

```{r}
black_actor_lead <- black_actor_lead %>% mutate(Year = as.numeric(Year))
latin_actor_lead <- latin_actor_lead %>% mutate(Year = as.numeric(Year))
asian_actor_lead <- asian_actor_lead %>% mutate(Year = as.numeric(Year))
egyptian_actor_lead <- egyptian_actor_lead %>% mutate(Year = as.numeric(Year))
```

```{r}
#adding Race variable
actor_lead$Race <- ifelse(actor_lead$Actor %in% black_actor_lead$Name, "Black", 
       ifelse(actor_lead$Actor %in% latin_actor_lead$Nominee, "Latin", 
              ifelse(actor_lead$Actor %in% asian_actor_lead$Name, "Asian",
                     ifelse(actor_lead$Actor %in% egyptian_actor_lead$Name, "Egyptian", "White"))))

#adding an id var to plot
actor_lead <- actor_lead %>% group_by(Year) %>% mutate(id = 1:n())

#adding decade groups
actor_lead$TimePeriod <- ifelse(actor_lead$Year <= 1950, "1928-1950",
              ifelse(actor_lead$Year <= 1970, "1951-1970",
                     ifelse(actor_lead$Year <= 1990, "1971-1990",
                            ifelse(actor_lead$Year <= 2010, "1991-2010", "2011-2021"))))
```


```{r}
t1 <- actor_lead %>% filter(TimePeriod == "1928-1950") %>% 
  ggplot(aes(x = as.factor(Year), y = id, col = Race)) +
  geom_point(size = 5) +
  theme_light() +
  scale_color_manual(values = c("#E69F00", "#000000")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.title = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  axis.ticks.y=element_blank(), axis.text.y=element_blank(), axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.2)) +
  coord_fixed(ratio=1.5) +
  geom_text_repel(aes(label = ifelse(Actor %in% c(black_actor_lead$Name, asian_actor_lead$Name, latin_actor_lead$Nominee, egyptian_actor_lead$Name), Actor, "")), size = 3, max.overlaps = Inf) +
  ggtitle("Actors: 1928-1950")

t1
```

```{r}
t2 <- actor_lead %>% filter(TimePeriod == "1951-1970") %>% 
  ggplot(aes(x = as.factor(Year), y = id, col = Race, label=Actor)) +
  geom_point(size = 5) +
  theme_light() +
  scale_color_manual(values = c("#56B4E9", "#CC79A7", "#E69F00", "#000000")) + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.title = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.2)) +
  geom_text_repel(aes(label = ifelse(Actor %in% c(black_actor_lead$Name, asian_actor_lead$Name, latin_actor_lead$Nominee, egyptian_actor_lead$Name), Actor, "")), size = 3, max.overlaps = Inf) +
  ggtitle("Actors: 1951-1970")

t2
```

```{r}
t3 <- actor_lead %>% filter(TimePeriod == "1971-1990") %>% 
  ggplot(aes(x = as.factor(Year), y = id, col = Race, label=Actor)) +
  geom_point(size = 5) +
  theme_light() +
  scale_color_manual(values = c("#56B4E9", "#CC79A7", "#E69F00", "#000000")) + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.title = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.2)) +
  geom_text_repel(aes(label = ifelse(Actor %in% c(black_actor_lead$Name, asian_actor_lead$Name, latin_actor_lead$Nominee, egyptian_actor_lead$Name), Actor, "")), size = 3, max.overlaps = Inf) +
  ggtitle("Actors: 1971-1990")

t3
```


```{r}
t4 <- actor_lead %>% filter(TimePeriod == "1991-2010") %>% 
  ggplot(aes(x = as.factor(Year), y = id, col = Race, label=Actor)) +
  geom_point(size = 5) +
  theme_light() +
  scale_color_manual(values = c("#56B4E9", "#CC79A7", "#000000")) + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.title = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.2)) +
  geom_text_repel(aes(label = ifelse(Actor %in% c(black_actor_lead$Name, asian_actor_lead$Name, latin_actor_lead$Nominee, egyptian_actor_lead$Name), Actor, "")), size = 3, max.overlaps = Inf) +
  ggtitle("Actors: 1991-2010")

t4
```


```{r}
t5 <- actor_lead %>% filter(TimePeriod == "2011-2021") %>% 
  ggplot(aes(x = as.factor(Year), y = id, col = Race, label=Actor)) +
  geom_point(size = 5) +
  theme_light() +
  scale_color_manual(values = c("#56B4E9", "#CC79A7", "#33a02c", "#E69F00", "black")) + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.title = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.2)) +
  geom_text_repel(aes(label = ifelse(Actor %in% c(black_actor_lead$Name, asian_actor_lead$Name, latin_actor_lead$Nominee, egyptian_actor_lead$Name), Actor, "")), size = 3, position = "dodge") +
  ggtitle("Actors: 2011-2021")

t5
```



